---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# covid19nytimes

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/covid19nytimes)](https://CRAN.R-project.org/package=covid19nytimes)
<!-- badges: end -->

The  covid19nytimes package harvests the data made freely available by the New York Times. See https://www.nytimes.com/article/coronavirus-county-data-us.html for more.

## Installation

You can install the released version of covid19nytimes from [CRAN](https://CRAN.R-project.org) with:

``` {r install, eval=FALSE}
install.packages("covid19nytimes")
```

Or the latest development version from [github](https://github.com/covid19R/covid19nytimes)

```{r install_github, eval = FALSE}
devtools::install_github("covid19R/covid19nytimes")
```

## Data

The package has the data from states and counties. The package comes with static data that was downloaded at the time of the last package update.

```{r data}
library(covid19nytimes)

head(covid19nytimes_states) %>% knitr::kable()

head(covid19nytimes_counties) %>% knitr::kable()

```

## Getting the Most Up to Date Data

To get the most updated data, run the following functions

```{r refresh eval=FALSE}
covid19nytimes_states <- refresh_covid19nytimes_states()

covid19nytimes_counties <- refresh_covid19nytimes_counties()
```

## Columns

The data follows the covid19R standard for tidy Covid-19 data. The data columns are as follows:

* date - The date in YYYY-MM-DD form
* location - The name of the location as provided by the data source. The counties dataset provides county and state. They are combined and separated by a `,`, and can be split by `tidyr::separate()`, if you wish.
* location_type - The type of location using the covid19R controlled vocabulary. Nested locations are indicated by multiple location types being combined with a `_
* location_standardized - A standardized location code using a national or international standard. In this case, FIPS state or county codes. See https://en.wikipedia.org/wiki/Federal_Information_Processing_Standard_state_code and https://en.wikipedia.org/wiki/FIPS_county_code for more
* location_standardized_type The type of standardized location code being used according to the covid19R controlled vocabulary. Here we use `fips_code`
* data_type - the type of data in that given row. Includes `total_cases` and `total_deaths`, cummulative measures of both.
* value - number of cases of each data type

## Sample visualization

```{r example_plot}
library(dplyr)
library(ggplot2)

covid19nytimes_states %>%
  filter(location %in% c("Washington", "New York", "Massachusetts", "Michigan", "Illinois")) %>%
  filter(type == "total_deaths") %>%
  ggplot(aes(x = date, y = value, color = location)) +
  geom_line() +
  theme_minimal() +
  scale_y_continuous(trans = "log10")
```
