---
title: "Making the NY Times Bubble Map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Making the NY Times Bubble Map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The NY Times has been making an amazing [infographic of Covid-19 cases](https://www.nytimes.com/interactive/2020/us/coronavirus-us-cases.html). With the data from this package, we should be able to more or less recreate it. Let's do it for the lower 48 (moving Hawaii and Alaska is a pain!)

We'll start by loading the coronavirus data at the county and state level
```{r setup, message=FALSE}
require(covid19nytimes)
require(dplyr)

covid19nytimes_states <- refresh_covid19nytimes_states()
covid19nytimes_counties <- refresh_covid19nytimes_counties()
```

Next, we need some county maps and state maps. For this, we'll use the excellent [USAboundaries](https://cran.r-project.org/web/packages/USAboundaries/index.html) package. We want the centroid of counties for bubble placement.

```{r get_counties, warning=FALSE, message=FALSE}
require(sf)
require(USAboundaries)

county_map <- us_counties() %>%
  st_centroid()

state_map <- us_states()

```

OK, let's filter to the lower 48 for all data sets

```{r filter}
filter_out <- c("Alaska", "Hawaii", "Guam", "Puerto Rico")

covid19nytimes_states <- covid19nytimes_states %>%
  filter(!(location %in% filter_out)) %>%
  filter(location != "Unknown")

covid19nytimes_counties <- covid19nytimes_counties %>%
  filter(purrr::map_dbl(location, ~stringr::str_detect(., filter_out) %>% sum)==0) %>%
  filter(location != "Unknown")

county_map <- county_map %>%
    filter(!(state_name %in% filter_out))

state_map <- state_map %>%
    filter(!(name %in% filter_out))
```

For text labels, we use centroids of states joined with maximum number of case numbers.

```{r case_num, warning=FALSE}
state_num_centroids <- covid19nytimes_states %>%
  filter(date == max(date)) %>%
  filter(data_type == "cases_total") %>%
  left_join(state_map,
            by = c("location_standardized" = "geoid")) %>%
  st_as_sf(crs = st_crs(state_map)) %>%
  st_centroid()
```

Now, we join the county data and the county centroid map and filter to the last date in the data set.

```{r join_counties}
joined_covid_counties <- covid19nytimes_counties %>%
  filter(date == max(date)) %>%
  filter(data_type == "cases_total") %>%
  left_join(county_map,
            by = c("location_standardized" = "geoid")) %>%
  st_as_sf(crs = st_crs(county_map))
```

Last, let's make the plot!

```{r plot, warning=FALSE, fig.width = 10, fig.height = 8}
require(ggplot2)

ggplot() +
  geom_sf(data = state_map, fill = "lightgrey", color = "darkgrey") +
  geom_sf_text(data = state_num_centroids, aes(label = value), color = "red",
                size = 4, alpha = 0.7) +
  geom_sf(data = joined_covid_counties,
          mapping = aes(size = value),
          shape = 21, alpha = 0.5, color = "red", fill = "pink") +
  theme_void() +
  scale_size_continuous(guide= "none",
                        range = c(0.05, 35)) +
  coord_sf(crs = st_crs(2163)) #for that nice equal area curved look
```
